---
title: "Kimbe Predator Fish Abundance"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

# Introduction
This document analyses predator fish abundance from the Kimbe Bay video survey data set.
Last edit: 30 Aug 2021 BJC

# Packages and housekeeping
```{r packages, include=FALSE, eval=TRUE}
library(gridExtra) #for plotting
library(writexl)
library(gt)
library(lme4)      #basic model builder
library(car)       #for regression diagnostics
library(multcomp)  #for lm work
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(DHARMa)    #for residual diagnostics
library(performance)
library(glmmTMB)   #for wider selection of error families
library(sjPlot)    #for outputs
library(effects)   #for partial effects plots
library(ggeffects) #for partial effects plots
library(emmeans)   #for estimating marginal means
library(modelr)    #for auxiliary modeling functions
library(artyfarty) #for plotting
library(tidyverse) #for data wrangling
```


# Data import and wrangling

## Rmd setup, directory selection (Note: you need to run this code in knitr not in isolation otherwise won't work)
```{r setup, include=TRUE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = "~/OneDrive - James Cook University/Ben PhD/Data & analysis/KimbePreds/code", echo = TRUE)
getwd()
```

## Basic data wrangling: Loads fish DOV data from both survey periods (2018 and 2019) and extracts just predator observations:
```{r diw-chunk, include=FALSE}
source("../pred_diw.R", local = knitr::knit_global())
```


## Basic abundance calculations on the predator data frame
```{r diw-abun-chunk, include=FALSE}
source("pred_abun_diw.R", local = knitr::knit_global())
```


## Variables:
Modelling predator abundance (response variable) against reef type (fixed effect predictor variable), with site as a random effect:

1. Numbers of predators (response variable): This is count data and needs to be scaled against total fish assemblage size, so could convert into a proportion (0-1) but instead we'll scale inside the model using offset().
2. Reef type (predictor variable (fixed effect)): Factor with 3 'levels'.
3. Site (nested random effect): 4 sites per reef type, 12 total.
4. Numbers of fish (scaling variable): Need to scale the model to account for total fish numbers

Note: Initially thought about including survey period as a fully crossed random effect, however there are only two levels within this variable making any variance partitioning meaningless (can't have variance between 2 observations - generally considered that 5 is the minimum). In theory could include as a fixed effect, however we are not really interested in it as a predictor and want to maintain model parsimony!


## Basic data exploration
Using predsum, check how the data are distributed across the different reefs and reef types
```{r abun-eda}
# Direct from script
#source("code/2.pred_abun_eda.R")
#source("code/2.pred_abun_eda.R", local = knitr::knit_global())

# Or:

# Histogram 
hist <- ggplot(predsum, aes(x=No_Preds)) + 
  scale_y_continuous(limits = c(0,120), "Observations") +
  #scale_x_binned("No. of predators")+
  geom_histogram(binwidth=50, color="black", fill="grey")

# Compare sites:
(predreeftypestats <- predsum %>% 
  group_by(Reeftype) %>% 
  summarise_each(funs(mean,sd,se=sd(.)/sqrt(n())), No_Preds))

rtbar <- ggplot(predreeftypestats, aes(x=Reeftype, y=mean))+
  geom_bar(stat="identity", position="dodge", color="black")+
  scale_y_continuous(limits = c(0,250), "Mean predator count (±SE)")+
  scale_x_discrete("Reef type and survey period")+
  labs(title="Predator abundance by reef type")+
  scale_fill_viridis_d()+
  theme_scientific()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))+
  theme(legend.key.size = unit(0.4, "cm"))+
  theme(plot.title = element_text(size = 14, face = "bold"),
        legend.title=element_text(size=9), 
        legend.text=element_text(size=8),
        axis.title.y = element_text(size=10, face="bold"),
        axis.title.x = element_text(size=10, face="bold"),
        axis.text = element_text(size=6))
grid.arrange(hist, rtbar, ncol=2)
```
- Lots of zeros so perhaps will need a ZI term in the model. These are count data so expecting Poisson or neg binom distributions to work best.
- Predator abundance high on pinnacles where we exclusively recorded the large schools of pelagic/semi-pelagic fishes (in particular S. qenie). Could look at transforming or adjusting the data, treating these as outliers and excluding.  However, these observations may be ecologically important and the right GLMM error-family should handle these data.  

## Fit models
Basic GLMM with site as random factor. Will introduce offset later on.
Trialled other error families but all poor fit: Gaussian, Poisson, Gamma, Beta (on %) etc

# Model formulas (have gone back and checked possible influence of survey period as a nested random effect)
```{r abun-model-fit}
# Neg binom 2 (Neg binom 1 poor fit)
pred.glmmnb <- glmmTMB(No_Preds~Reeftype + (1|SiteCode), data=predsum, family=nbinom2())  
pred.glmmnb1 <- glmmTMB(No_Preds~Reeftype + (1|SurvCode/SiteCode), data=predsum, family=nbinom2())  
pred.glmmnb2 <- glmmTMB(No_Preds~Reeftype + (1|SiteCode/SurvCode), data=predsum, family=nbinom2())  
pred.glmmnb3 <- glmmTMB(No_Preds~Reeftype + (1|SiteSurv), data=predsum, family=nbinom2())  
pred.glmmnb4 <- glmmTMB(No_Preds~Reeftype + SurvCode + (1|SiteCode), data=predsum, family=nbinom2())  


# ZI (on neg binom 2)
pred.glmmzi <- glmmTMB(No_Preds~Reeftype+(1|SurvCode/SiteCode), data=predsum, family=nbinom2(), ziformula=~1)
# Tweedie
pred.glmmt <- glmmTMB(No_Preds~Reeftype+(1|SurvCode/SiteCode), data=predsum, family=tweedie())  
```

# Source code for all models trialled if required (model comparisons etc)
```{r abun-model-source}
# source("3a.pred_abun_model.R", local = knitr::knit_global())
```

##  Model diagnostics 

# DHARMa residuals:
```{r abun-model-diag}
nb.resid <- simulateResiduals(pred.glmmnb, plot=T) # 
zi.resid <- simulateResiduals(pred.glmmzi, plot=T) #
t.resid <- simulateResiduals(pred.glmmt, plot=T) # 
```
-> All look ok apart from overdispersion

# All NB models together
```{r}
simulateResiduals(pred.glmmnb, plot=T)
simulateResiduals(pred.glmmnb1, plot=T)
simulateResiduals(pred.glmmnb2, plot=T)
simulateResiduals(pred.glmmnb3, plot=T)
simulateResiduals(pred.glmmnb4, plot=T)
```



# Validation with AIC
```{r abun-aic1}
AIC(pred.glmmnb, pred.glmmzi, pred.glmmt)
```
-> NB model best and most parsimonious fit, but will check for zero inflation on that model

# All NB models together
```{r}
AIC(pred.glmmnb, pred.glmmnb1, pred.glmmnb2, pred.glmmnb3, pred.glmmnb4)
```


# ZI interrogation
```{r abun-nb-zicheck}
testZeroInflation(pred.glmmnb)
```
# ZI test output
data:  simulationOutput
ratioObsSim = 0.73892, p-value = 0.24
alternative hypothesis: two.sided
-> Plot shows that we are actually getting slightly fewer zeros than expected and the output quantifies this: 74% of zeros expected (non significant)

# Can also interrogate the ZI model using summary()
```{r zi-sum}
summary(pred.glmmzi)
```
# Manual calculation of probability and/or odds of a false zero
## Use plogis to transform the ZI estimate:
plogis(-21.2)= 6.2e-10 -> almost no chance that any of these zeros are false
exp(-21.2) -> and corresponding almost no odds that any zeros false either


# Model interrogation
```{r}
summary(pred.glmmnb)
```


```{r}
summary(pred.glmmnb1)
```


```{r}
summary(pred.glmmnb2)
```


```{r}
summary(pred.glmmnb3)
```
# With SurvCode as Fixed Effect
```{r}
summary(pred.glmmnb4)
```

```{r}

```


```{r abun-nb-eda}
emmeans(pred.glmmnb, pairwise~Reeftype, transform="response") %>% 
  confint()
emmeans(pred.glmmnb2, pairwise~Reeftype, transform="response") %>% 
  confint()
emmeans(pred.glmmnb4, pairwise~SurvCode, transform="response") %>% 
  confint()
```
-> So emmeans suggest that pinnacles have more predators than both offshore emergent and near shore reefs, but that this is highly variable.  What about if we take into account overall abundance for whole fish communities?


# Adjust model to account for variations in overall fish communities
-> On reefs where there are just generally no fish there will be consequently lower predators
We could analyse predator abundance as a proportion/% of overall fish abundance but modelling proportions has its own issues around error distributions (a defined scale 0-100% where variance decreases towards either end of the scale)
So we can add an offset instead
On log() scale as that is the link scale
```{r abun-nb-offset, message=FALSE, }
pred.glmmnboff <- update(pred.glmmnb, ~ . + offset(log(No_Fish)))
pred.glmmnboff2 <- update(pred.glmmnb2, ~ . + offset(log(No_Fish)))

```


# Revalidation with AIC
```{r abun-aic2}
AIC(pred.glmmnb,  pred.glmmnboff, pred.glmmnboff2)
```
-> Improved fit for offset model


# Check summary
```{r abun-model-summary}
summary(pred.glmmnboff)
```

```{r}
emmeans(pred.glmmnboff, pairwise ~ Reeftype, transform='response', combine=TRUE) %>%
  confint()
```


# Results (emmeans and contrasts on response scale - numbers of predators)
```{r abun-emmeans-results}
# emmeans and contrasts first
abuncontrasts <- 
  emmeans(pred.glmmnboff, pairwise ~ Reeftype, transform='response', combine=TRUE) %>%
  confint() %>% 
  as.data.frame() 

# pvalues if required/interested
abunpvals <- 
  emmeans(pred.glmmnboff,  ~ Reeftype, transform='response', combine=TRUE) %>% 
  pairs() %>% 
  as.data.frame() 
  
# Combine for full results table
abunresults <- 
  bind_cols(abuncontrasts, abunpvals[5:6]) 

# Write to csv
write_csv(abunresults, file = "../../output/rtables/abunresults.csv")

# Copy to html
abunresults %>% 
  tab_df(file="../../output/rtables/abunresults.html", title = "Abundance contrasts (estimates)") #%>% 

# Display inline
abunresults
```





# Can also check contrast ratios (using type="response"):
```{r abun-emmeans-ratios}
# Ratios with CI
abunratios <- 
  emmeans(pred.glmmnboff, pairwise ~ Reeftype, type='response', combine=TRUE) %>%
  confint() %>% 
  as.data.frame()

# p values
abunratiopvals <- emmeans(pred.glmmnboff,  ~ Reeftype, type='response', combine=TRUE) %>% 
  pairs() %>% 
  as.data.frame() 

# Combine
abunratios <- bind_cols(abunratios, abunratiopvals[5:6]) %>% 
  mutate_if(is.numeric, round, 2)

# Save as RData for plotting
save(abunratios, file='../../data/abunratios.RData')


# Write to csv
write_csv(abunratios, file = "../../output/rtables/abunratios.csv")

# Copy to html 
abunratios %>% 
  tab_df(file="../../output/rtables/abunratios.html", title = "Abundance contrasts (ratios)")
abunratios

```

# Create plot data
Final step is to create the df to be used for plot making 
```{r abun-plot-data}
# First calculate the emmeans estimates
abunplotmeans <- emmeans(pred.glmmnboff, pairwise ~ Reeftype, transform='response') %>% 
  confint() %>% 
  as_tibble() %>% 
  #dplyr::select(-contrasts) %>% 
  transmute(Reeftype = emmeans$Reeftype,
            Mean = emmeans$response,
            DF = emmeans$df,
            MeanLowerCI95 = emmeans$lower.CL,
            MeanUpperCI95 = emmeans$upper.CL
            )


# And bind to the 95% CI contrasts 
abunplotcontrasts <- 
  pred.glmmnboff %>% 
  emmeans(pairwise ~ Reeftype, transform='response') %>%
  confint() %>% 
  as_tibble() %>% 
  transmute(
    Contrast = contrasts$contrast, 
    Estimate = contrasts$estimate, 
    contrastlowerCI95 = contrasts$lower.CL,
    contrastupperCI95 = contrasts$upper.CL)         

(abunplotdata <- abunplotmeans %>% 
  bind_cols(abunplotcontrasts) )

# Save as RData for plotting
save(abunplotdata, file='../../data/abunplotdata.RData')

# HTML
abunplotdata %>% 
  tab_df(file="../../output/rtables/abunresults.html", title = "Abundance results")
```

..And table for publication
```{r abun-pub-data}
# Write to excel for pasting into publication - need to rename some of the columns etc.
abunpubdata <- 
abunplotdata %>% 
  mutate_if(is.numeric, round, 1) %>% 
  bind_cols(abunratios[c(8,11,12)]) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  mutate(
    `Abundance (n.150m-2)` = Mean,
    `EMM 95% CI`= paste(MeanLowerCI95, MeanUpperCI95, sep = '-'), 
    .before = DF, .keep = 'unused') %>% 
  mutate(
    `Pairwise contrasts` = Contrast, 
    `Effect size (n.150m-2)` = Estimate,
    `Effect size 95% CI` = paste(contrastlowerCI95, contrastupperCI95, sep = '-'), 
     .keep = 'unused') %>% 
  mutate(
    `Contrast ratio` = contrasts.ratio,
    `Contrast ratio 95% CI` = paste(contrasts.lower.CL, contrasts.upper.CL, sep = '-'),
    .keep = 'unused')

abunpubdata
write_xlsx(abunpubdata, path = "../../output/rtables/abunpubdata.xlsx")
```





```{r run-whole-script-chunk, include=TRUE, warnings=FALSE, message=FALSE}
# Run this chunk to execute whole script without producing html output
```





