---
title: "Preds NMDS work - all 0 rows left in, fraction added"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  code_folding: show
  collapse: no
  df_print: paged
  fig_caption: yes
  fig_height: 4
  fig_width: 4
  highlight: textmate
  theme: spacelab
  toc: yes
  toc_float: yes
editor_options: 
  chunk_output_type: inline
---

# Created 22 Nov 2021 BJC - following discussions with HBH that NMDS is the preferable ordination technique for these data
Inspiration from https://ourcodingclub.github.io/tutorials/ordination/#section5



# Preliminaries {.tabset .tabset-faded}

## Set up
```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
# Rmd setup, directory selection
knitr::opts_chunk$set(root.dir = "~/OneDrive - James Cook University/Ben PhD/Data & analysis/KimbePreds/code/PredDiversity")
getwd()
```


## Housekeeping
```{r housekeep}
rm(list=ls())
graphics.off()
```


## Packages
```{r packages}
library(MASS)
library(ggdendro)
library(vegan)
library(ggvegan)
library(tidyverse)
```

# Wrangle in data {.tabset .tabset-faded}

## Load
Start with predator diversity df - preddiv
```{r data-load}
load(file='../../data/preddiv.RData')
```

## Check
```{r data-checks}
head(preddiv)
```

First column is TID, followed by grouping/environmental descriptors and diversity metrics etc.
The species observations start at column 11
We'll need to separate all these out (metaMDS only works on a matrix of spp observations)


## Create env
- Pull out the descriptors into an object called env and add in reef type
```{r create-env}
env <- preddiv[1:10] %>%
  mutate(Reeftype = factor(case_when(grepl("BRAD", TID) ~ "Pinnacle",
                                     grepl("JOEL", TID) ~ "Pinnacle",
                                     grepl("KBOM", TID) ~ "Pinnacle",
                                     grepl("INGL", TID) ~ "Pinnacle",
                                     grepl("LADI", TID) ~ "Nearshore",
                                     grepl("MADA", TID) ~ "Nearshore",
                                     grepl("SUSA", TID) ~ "Nearshore",
                                     grepl("DON", TID) ~ "Nearshore",
                                     grepl("EMA", TID) ~ "Offshore",
                                     grepl("HOG", TID) ~ "Offshore",
                                     grepl("OTT", TID) ~ "Offshore",
                                     grepl("KIS", TID) ~ "Offshore")))
```


## Spp obs
- Create matrix with only spp observations
```{r spp-matrix}
predmatrix <- preddiv[c(1,11:73)] %>%  # Keep TID in
  column_to_rownames("TID") %>%        # And then set as row names
    +0.01 #%>%
  #filter_all(any_vars(. != 0)) # Drop transects with only 0 observations (22 in total)
```

- and convert to distance matrix, based on Bray-Curtis
```{r distance-matrix}
preddist <- vegdist(predmatrix,  method = "bray")
```

# Create NMDS:

Run mds on preddist
```{r mds1}
predNMDS <- metaMDS(preddist, k = 2, trymax = 100, trace = FALSE) 
# Check
predNMDS # Works - Stress 0.15
```

## Stressplot
```{r}
stressplot(predNMDS)
```

## Basic plot
### Pull out scores etc
```{r mds2}
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
```

### And plot
```{r mds-plot1}
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()

```



# Investigation

## Extract species scores
```{r}
species.scores <- as.data.frame(scores(predNMDS, "species"))  # Using the scores function from vegan to extract the species scores and convert to a data.frame
```
-> Doesn't work!! No species scores included as we ran this on the distance object we created from spp observations, which dropped the species from the object.

But we can run the mds again on predmatrix rather than preddist as long as autotransform is turned off:

```{r}
set.seed(2)
predNMDS <- metaMDS(predmatrix, k = 2, trymax = 100, trace = FALSE, autotransform = FALSE, distance="bray")
```

## Check that spp scores can be extracted
```{r}
as.data.frame(scores(predNMDS, "species")) %>% 
  head
```

Now spp scores can be extracted, along with site etc.


# Extract data scores: {.tabset .tabset-faded}

## Species scores
```{r}
species.scores <- as.data.frame(scores(predNMDS, "species")) %>% 
  rownames_to_column(var = "species")
```

And need to filter out the most important ones:
```{r}

species.scores <- species.scores %>% 
  filter(species %in% c("Sphyraena qenie", "Caranx sexfasciatus", "Macolor macularis",
                        "Caranx melampygus", "Lutjanus gibbus", "Cephalopholis cyanostigma"))


```


## Adjust for plotting
And need to make separate columns with 'coords' for the labels, so that they dodge the other information on the final graph

1. C mel
2. C sex
3. L gib
4. M mac
5. C cya
6. S qen 

```{r}
#             1      2       3       4       5        6
Coord_1 <- c(-2.6,	-4.2,	 -2.1,	  -2.6,	   1.2,	     -6.5)
Coord_2 <- c(-0.2,	0.2,	   0.9,	   0.5,	  -0.1,	   -0.7)


species.labels <- species.scores %>% 
  mutate(Coord_1 = Coord_1,
         Coord_2 = Coord_2)

```

## Site scores
```{r}
data.scores <- as.data.frame(scores(predNMDS))  # Scores function from vegan extracts the site scores -> data.frame
data.scores$TID <- rownames(data.scores)  # create a column of site names, from the rownames of data.scores
data.scores$grp <- env$Reeftype  #  going to want to group by reef type

shannon <- env %>% 
  dplyr::select("TID", "Shannon")



data.scores <- data.scores %>% 
  full_join(shannon, by = "TID")
#head(data.scores)  #look at the data
```

# Plot

```{r}
pred_nmds_plot <-
  ggplot(data.scores, aes(NMDS1, NMDS2)) + # Each individual TID score in NMDS space
  geom_point(aes(colour = grp, shape = grp), alpha=0.6, position=position_jitter(.1)) +  # grouping by reef type (grp)
  # geom_point(aes(colour = grp, shape = grp, size=Shannon), alpha=0.6, position=position_jitter(.1)) +  # or could even include scale by Shannon score
  xlim(-8,2.5)+
  ylim(-3, 3)+
  scale_color_manual(values=c("#DFC27D", "#436EEE", "#35978F"))+ # The classic Kimbe colour scheme
  scale_fill_manual(values=c("#DFC27D", "#436EEE", "#35978F"))+ # The classic Kimbe colour scheme
  scale_shape_manual(values=c(16, 8, 17))+                       # and classic Kimbe pch scheme
  stat_ellipse(geom = "polygon", aes(colour = grp, fill = grp), alpha=0.1, type='t',size =0.8) +          # 95% conf ellipses
  stat_ellipse(aes(colour = grp), alpha=1, type='t',size =0.8) +          # 95% conf ellipses
  geom_point(data = species.scores, aes(NMDS1, NMDS2), pch = 15, size = 2,
             position=position_jitter(0.1)) +                    # Pts for Spp scores for top 6 taxa
  geom_text(data = species.labels, aes(Coord_1, Coord_2, label = species), size = 4, fontface = "italic") + # Text for these           
  #geom_segment(data = species.scores, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2),                # And arrows
    #           arrow = arrow(length = unit(0.2, 'cm')), colour = "black", alpha=0.6)+
  #geom_text(aes(label = TID), size = 2)+  
  geom_hline(yintercept = 0, linetype=2, alpha=0.3) +     # 0 on y axis
  geom_vline(xintercept = 0, linetype=2, alpha=0.3) +     # 0 on x axis
  theme_classic()+                        # White fill
  theme(
    axis.line = element_blank(), 
    panel.border = element_rect(size = 1, fill = "transparent"),
    legend.title = element_blank(),
    legend.text = element_text(face = "plain", size = 10),
    legend.position = c(.1,.13),
    legend.background = element_rect(fill = "transparent"),
    legend.box.background = element_rect(colour = "black"),
    legend.spacing.y = unit(0.1, "mm"),
    legend.key.size = unit(5, "mm")
    )

pred_nmds_plot

#ggsave(pred_nmds_plot, filename= '../../output/rfigures/prednmds.pdf',  width = 174, height = 109, units = "mm", dpi = 600)
```



## Add the venn diagram in as a grob file (see separate script for venn diag creation)

```{r}
nmdsvenn <- 
pred_nmds_plot + annotation_custom(
    ggplotGrob(predvenn), 
    xmin = -13, xmax =0, ymin = -0.2, ymax = 3.2)

nmdsvenn
ggsave(nmdsvenn, filename= '../../output/rfigures/nmdsvenn.png',  width = 174, height = 109, units = "mm", dpi = 1000 )

```









