---
title: "Preds NMDS work"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
  code_folding: show
  collapse: no
  df_print: paged
  fig_caption: yes
  fig_height: 4
  fig_width: 4
  highlight: textmate
  theme: spacelab
  toc: yes
  toc_float: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Created 22 Nov 2021 BJC - following discussions with HBH that NMDS is the preferable ordination technique for these data

Inspiration from https://ourcodingclub.github.io/tutorials/ordination/#section5

# Preliminaries {.tabset .tabset-faded}

## Set up
```{r setup, include=FALSE, warnings=FALSE, message=FALSE}
# Rmd setup, directory selection
knitr::opts_chunk$set(root.dir = "~/OneDrive - James Cook University/Ben PhD/Data & analysis/KimbePreds/code/PredDiversity", echo = TRUE)
getwd()
```


## Housekeeping
```{r housekeep}
rm(list=ls())
graphics.off()
```

## Packages
```{r packages}
library(MASS)
library(ggdendro)
library(vegan)
library(ggvegan)
library(tidyverse)
```

# Wrangle in data {.tabset .tabset-faded}

## Load
Start with predator diversity df - preddiv
```{r data-load}
load(file='../../data/preddiv.RData')
```

## Check
```{r data-checks}
head(preddiv) # First column is TID and the spp. observations start at column 11, so we'll need to separate these out
```

## Create env
- Pull out the descriptors into an object called env and add in reef type
```{r create-env}
env <- preddiv[1:10] %>%
  mutate(Reeftype = factor(case_when(grepl("BRAD", TID) ~ "Pinnacle",
                                     grepl("JOEL", TID) ~ "Pinnacle",
                                     grepl("KBOM", TID) ~ "Pinnacle",
                                     grepl("INGL", TID) ~ "Pinnacle",
                                     grepl("LADI", TID) ~ "Nearshore",
                                     grepl("MADA", TID) ~ "Nearshore",
                                     grepl("SUSA", TID) ~ "Nearshore",
                                     grepl("DON", TID) ~ "Nearshore",
                                     grepl("EMA", TID) ~ "Offshore",
                                     grepl("HOG", TID) ~ "Offshore",
                                     grepl("OTT", TID) ~ "Offshore",
                                     grepl("KIS", TID) ~ "Offshore")))
```



## Spp obs
- Create matrix with only spp observations
```{r spp-matrix}
predmatrix <- preddiv[c(1,11:71)] %>% 
  column_to_rownames("TID") %>%
  filter_all(any_vars(. != 0)) # Drop transects with only 0 observations (22 in total)
```

- and convert to distance matrix, based on Bray-Curtis
```{r distance-matrix}
preddist <- vegdist(predmatrix,  method = "bray")
```

However, this won't work with metaMDS
```{r mds1}
predNMDS <- metaMDS(preddist, k = 2, trymax = 100, trace = FALSE)
```

### So probably going to need some sort of transformation or will need to add a small fraction to every score





# Transformations {.tabset .tabset-faded}

### Decostand: A function for implementing many of the most effective/popular standardisations in community ecology

## total: 
- divide by margin total (default MARGIN = 1).
```{r trans-total}
preddist_tot <- decostand(preddist, method = "total")
predNMDS <- metaMDS(preddist_tot, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```



## max: 
- divide by margin maximum (default MARGIN = 2).
```{r trans-max}
preddist_max <- decostand(preddist, method = "max")
predNMDS <- metaMDS(preddist_max, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```

## frequency: 
- divide by margin total and multiply by the number of non-0 items, so that avg of non-0 entries is one (Oksanen 1983; default MARGIN = 2).
```{r trans-freq}
preddist_freq <- decostand(preddist, method = "frequency")
predNMDS <- metaMDS(preddist_freq, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```

## normalize: 
- make margin sum of squares equal to one (default MARGIN = 1).
```{r trans-norm}
preddist_norm <- decostand(preddist, method = "normalize")
predNMDS <- metaMDS(preddist_norm, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```


## range: 
- standardize values into range 0 ... 1 (default MARGIN = 2). If all values are constant, they will be transformed to 0.
```{r trans-range}
preddist_range <- decostand(preddist, method = "range")
predNMDS <- metaMDS(preddist_range, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```

## rank, rrank: 
- rank replaces abundance values by their increasing ranks leaving zeros unchanged, and rrank is similar but uses relative ranks with maximum 1 (default MARGIN = 1). Average ranks are used for tied values.
```{r trans-rank}
preddist_rank <- decostand(preddist, method = "rrank")
predNMDS <- metaMDS(preddist_rank, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```


## standardize: 
- scale x to zero mean and unit variance (default MARGIN = 2).
```{r trans-stdz}
preddist_std <- decostand(preddist, method = "standardize")
predNMDS <- metaMDS(preddist_std, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```

## pa: 
- scale x to presence/absence scale (0/1).
```{r trans-pa}
preddist_pa <- decostand(preddist, method = "pa")
predNMDS <- metaMDS(preddist_pa, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```

## chi.square:
- divide by row sums and square root of column sums, and adjust for square root of matrix total (Legendre & Gallagher 2001).
```{r trans-chi2}
preddist_chi2 <- decostand(preddist, method = "chi.square")
predNMDS <- metaMDS(preddist_chi2, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```

## hellinger: 
- square root of method = "total" (Legendre & Gallagher 2001).
```{r trans-helling}
preddist_helling <- decostand(preddist, method = "hellinger")
predNMDS <- metaMDS(preddist_helling, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```

## log: 
- logarithmic transformation as suggested by Anderson et al. (2006): log_b (x) + 1 for x > 0, where b is the base of the logarithm; zeros are left as zeros. Higher bases give less weight to quantities and more to presences, and logbase = Inf gives the presence/absence scaling. Please note this is not log(x+1). Anderson et al. (2006) suggested this for their (strongly) modified Gower distance (implemented as method = "altGower" in vegdist), but the standardization can be used independently of distance indices.
```{r trans-log}
preddist_log <- decostand(preddist, method = "log")
predNMDS <- metaMDS(preddist_log, k = 2, trymax = 100, trace = F)
NMDS1 <- predNMDS$points[,1] ##also found using scores(predNMDS)
NMDS2 <- predNMDS$points[,2]
pred.plot<-cbind(predmatrix, NMDS1, NMDS2) %>% 
  rownames_to_column(var = "TID") %>% 
  left_join(env, by = "TID", keep = FALSE)
ggplot(pred.plot, aes(NMDS1, NMDS2, color=Reeftype))+
  geom_point(position=position_jitter(.1), shape=3)+##separates overlapping points
  #geom_text(aes(label = TID))+
  stat_ellipse(type='t',size =1)+ ##draws 95% confidence interval ellipses
  theme_minimal()
```









```{r}
predmatrix <- preddiv[c(1,11:71)] %>% 
  column_to_rownames("TID") %>%
  filter_all(any_vars(. != 0)) %>% # Drop 22 transects with only 0 observations
  +0.00001
```




