---
title: "KimbePredsBiomass"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
---

# Rmd created: 2.7.2021 BJC
# Last edit: 31.8.2021 BJC on Argo

## Introduction
This document analyses predator fish biomass from the Kimbe Bay video survey data set. 

## Packages and housekeeping
```{r biom-packages, include=FALSE, eval=TRUE}
#rm(list=ls())
#graphics.off()
#getwd()
library(gridExtra) #for plotting
library(writexl)
library(gt)
library(lme4)      #basic model builder
library(car)       #for regression diagnostics
library(multcomp)  #for lm work
library(broom)     #for tidy output
library(ggfortify) #for model diagnostics
library(DHARMa)    #for residual diagnostics
library(performance)
library(glmmTMB)   #for wider selection of error families
library(sjPlot)    #for outputs
library(effects)   #for partial effects plots
library(ggeffects) #for partial effects plots
library(emmeans)   #for estimating marginal means
library(modelr)    #for auxillary modelling functions
library(artyfarty) #for plotting
library(MuMIn)     #for pseudo R2
library(tidyverse) #for data wrangling
```

# Rmd setup, directory selection
```{r biom-setup, include=TRUE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = "code/PredBiomass", echo = TRUE)
#getwd()
```


# Basic data wrangling: Loads fish DOV data from both survey periods (2018 and 2019) and extracts just predator observations:
```{r diw-chunk, include=FALSE}
#source("../pred_diw.R", local = knitr::knit_global())
#Or can skip this as the next chunk includes importing main df
```


## Basic biomass calculations on the predator data frame
```{r biom-diw-chunk, results='markdown', eval=FALSE, message=FALSE, warning=FALSE}
source("pred_biom_diw.R", local = knitr::knit_global())
```

# Variables:
Modeling predator biomass (response variable) against reef type (fixed effect predictor variable), with site as a random effect:

1. Biomass (response variable - raw data in grams but diw converts to kg for rest of analysis)
2. Reef type (predictor variable - fixed effect): Factor with 3 'levels'.
3. Site (nested random effect): 4, nested within reef type, 12 total.
4. Total biomass of fish (scaling variable): Need to scale the model to account for total fish biomass?? --> Check if this makes a difference


# Basic data exploration
merpredsum has the transect_biomass column in it.
Can also run using: source("pred_biom_eda.R", local = knitr::knit_global())
```{r biom-eda1}
# Load data (if not running above chunks)
load(file = "../../data/merpredsum.RData")

# Glimpse
glimpse(merpredsum)

# Fig. 1: Histogram of all data 
hist1 <- ggplot(merpredsum, aes(x=Transect_pred_biomass_kg)) + 
  geom_histogram(color="black", fill="grey")

# Fig. 2: Histogram of data on log scale 
hist2 <- ggplot(merpredsum, aes(x=log(Transect_pred_biomass_kg))) + 
  geom_histogram(color="black", fill="grey")

# Table 1 and Fig. 3: Compare sites:
(predreeftypestats <- merpredsum %>% 
  group_by(Reeftype) %>% 
  summarise_each(funs(mean,sd,se=sd(.)/sqrt(n())), Transect_pred_biomass_kg))

# Fig 3
rtbar <- ggplot(predreeftypestats, aes(x=Reeftype, y=mean))+
  geom_bar(stat="identity", position="dodge", color="black")+
  scale_y_continuous("Mean predator count (±SE)")+
  scale_x_discrete("Reef type and survey period")+
  labs(title="Predator biomass by reef type")+
  scale_fill_viridis_d()+
  theme_scientific()+
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9))+
  theme(legend.key.size = unit(0.4, "cm"))+
  theme(plot.title = element_text(size = 14, face = "bold"),
        legend.title=element_text(size=9), 
        legend.text=element_text(size=8),
        axis.title.y = element_text(size=10, face="bold"),
        axis.title.x = element_text(size=10, face="bold"),
        axis.text = element_text(size=6))

# And plot
grid.arrange(hist1, hist2, rtbar, nrow=1)
```
- Lots of zeros so perhaps will need a ZI term in the model. 
- Log hist looks normal so prob will need a distribution on the log scale
- Not count data but expecting neg binom distributions to work best.
- As per abundance, predator biomass highest on pinnacles 
- Exaggerated by large schools of large pelagic/semi-pelagic fishes (in particular S. qenie)

# Fit models
Will start with basic GLMM with site as random factor. Definitely not going to be Gaussian, Gamma (zeros), Beta (not 0-1). 
Likely going to be a tweedie (a=can't really be Poisson, Neg Bin etc) but will include all of these for testing 
Will introduce offset later on to account for total fish biomass
```{r biom-model-fit, warning=FALSE}
# Gaussian - fitting anyways 
biom.glmm <- glmmTMB(Transect_pred_biomass_kg~Reeftype+(1|SiteCode), data=merpredsum)
# Poisson
biom.glmmp <- glmmTMB(Transect_pred_biomass_kg~Reeftype+(1|SiteCode), data=merpredsum, family=poisson())
# Negbinom
biom.glmmnb2 <- glmmTMB(Transect_pred_biomass_kg~Reeftype+(1|SiteCode), data=merpredsum, family=nbinom2(link = "log"))  
# ZI on nb
biom.glmmzi <- glmmTMB(Transect_pred_biomass_kg~Reeftype+(1|SiteCode), data=merpredsum, family=nbinom2(), ziformula=~1)
# Tweedie
biom.glmmt <- glmmTMB(Transect_pred_biomass_kg~Reeftype+(1|SiteCode), data=merpredsum, family=tweedie(link = "log"))  
```

# Validation with AIC
```{r biom-aic1}
AIC(biom.glmm, biom.glmmp, biom.glmmnb2, biom.glmmzi, biom.glmmt)
```

NB best AIC score but still need the usual diagnostic checks..


# Model diagnostics
```{r biom-model-diag}
simulateResiduals(biom.glmm, plot=T) # 
simulateResiduals(biom.glmmp, plot=T) # 
simulateResiduals(biom.glmmnb2, plot=T) 
simulateResiduals(biom.glmmzi, plot=T) 
simulateResiduals(biom.glmmt, plot=T)
```
- As expected Tweedie best residual fit although strictly speaking fails dispersion test. Still, we'll work with this family moving forward

# First will just check that the overdisp isn't coming from zi:
```{r zit}
biom.glmmt %>% testZeroInflation() # Looks good
```

# Without accounting for whole fish assemblage biomass
```{r biom-mod-emmeans}
emmeans(biom.glmmt, pairwise~Reeftype, transform="response") %>% 
  confint()
```


# Model offset (against total fish biomass)
Going to add the offset straight away
```{r biom-tweedie-offset}
biom.glmmtoff <- update(biom.glmmt, ~ . + offset(log(Transect_fish_biomass_kg)))
```

# Revalidation with AIC
```{r biom-aic2}
AIC(biom.glmmtoff, biom.glmmt)
```

- Offset model better fit


## Check non-offset model in emmeans first

```{r}
emmeans(biom.glmmt, pairwise ~ Reeftype, transform='response') %>%
  confint() %>% 
  as.data.frame() 


```




# Check summary
```{r}
summary(biom.glmmtoff)
```


# Results (emmeans and contrasts on response scale - predator biomass)
```{r biom-emmeans-results}
# emmeans and contrasts first
biomcontrasts <- 
  emmeans(biom.glmmtoff, pairwise ~ Reeftype, transform='response') %>%
  confint() %>% 
  as.data.frame() 

# pvalues if required/interested
biompvals <- 
  emmeans(biom.glmmtoff,  ~ Reeftype, transform='response', combine=TRUE) %>% 
  pairs() %>% 
  as.data.frame() 
  
# Combine for full results table
biomresults <- 
  bind_cols(biomcontrasts, biompvals[5:6]) 

# Write to csv
write_csv(biomresults, file = "../../output/rtables/biomresults.csv")

# Copy out to html
biomresults %>% 
 tab_df(file="../../output/rtables/biomresults.html", title = "Biomass (kg) contrasts (estimates)") #%>% 

# Display inline
biomresults
```
So non-significant difference between pinnacles and offshore reefs but other contrasts are significant

# Ratios (using type="response"):
```{r biom-emmeans-ratios}
# Ratios with CI
biomratios <- emmeans(biom.glmmtoff, pairwise ~ Reeftype, type='response', combine=TRUE) %>%
  confint() %>% 
  as.data.frame()
```


```{r biom-emmeans-ratios}
biom.glmmtoff %>% 
emmeans(pairwise ~ Reeftype, type='response', combine=TRUE) %>%
  confint() %>% 
  as.data.frame()

biom.glmmtoff %>% 
emmeans( ~ Reeftype, type='response',) %>% 
  pairs() %>% 
  summary(infer=TRUE) %>% 
  as.data.frame()

```


```{r biom-emmeans-ratios}
# p values
biomratiopvals <- emmeans(biom.glmmtoff,  ~ Reeftype, type='response', combine=TRUE) %>% 
  pairs() %>% 
  as.data.frame() 

# Combine
biomratios <- bind_cols(biomratios, biomratiopvals[5:6]) 

# Write to csv
write_csv(biomratios, file = "../../output/rtables/biomratios.csv")

# Copy to html
biomratios %>% 
  tab_df(file="../../output/rtables/biomratios.html", title = "Biomass (kg) contrasts (ratios)")

# Display inline
biomratios
```
CIs different with the ratios - something to do with the order of back-calculation?  The story is pretty much the same however.

# Create plot data
Final step is to create the df to be used for plot making 
```{r biom-plot-data}

# First calculate the emmeans estimates
biomplotmeans <- emmeans(biom.glmmtoff, pairwise ~ Reeftype, transform='response') %>% 
  confint() %>% 
  as_tibble() %>% 
  #dplyr::select(-contrasts) %>% 
  transmute(Reeftype = emmeans$Reeftype,
            Mean = emmeans$response,
            DF = emmeans$df,
            MeanLowerCI95 = emmeans$lower.CL,
            MeanUpperCI95 = emmeans$upper.CL
            )


# And bind to the 95% CI contrasts 
biomplotcontrasts <- emmeans(biom.glmmtoff, pairwise ~ Reeftype, transform='response') %>%
  confint() %>% 
  as_tibble() %>% 
  transmute(
    Contrast = contrasts$contrast, 
    Estimate = contrasts$estimate, 
    contrastlowerCI95 = contrasts$lower.CL,
    contrastupperCI95 = contrasts$upper.CL)

(biomplotdata <- biomplotmeans %>% 
  bind_cols(biomplotcontrasts))


save(biomplotdata, file='../../data/biomplotdata.RData')

biomplotdata %>% 
  tab_df(file="../../output/rtables/biomresults.html", title = "Biomass results")
```


```{r abun-pub-data}
# Write to excel for pasting into publication - need to rename some of the columns etc.
biompubdata <- 
biomplotdata %>% 
  mutate_if(is.numeric, round, 1) %>% 
  bind_cols(biomratios[c(8,11,12)]) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  mutate(
    `Biomass (kg.150m-2)` = Mean,
    `EMM 95% CI`= paste(MeanLowerCI95, MeanUpperCI95, sep = '-'), 
    .before = DF, .keep = 'unused') %>% 
  mutate(
    `Pairwise contrasts` = Contrast, 
    `Contrast effect size` = Estimate,
    `Contrast 95% CI` = paste(contrastlowerCI95, contrastupperCI95, sep = '-'), .keep = 'unused') %>% 
  mutate(
    `Contrast ratio` = contrasts.ratio,
    `Contrast ratio 95% CI` = paste(contrasts.lower.CL, contrasts.upper.CL, sep = '-'),
    .keep = 'unused')

biompubdata
write_xlsx(biompubdata, path = "../../output/rtables/biompubdata.xlsx")
```









