---
title: "Kimbe Predator Graphics - Supplementary Information"
author: "Ben Cresswell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
 html_document:
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: textmate
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: inline
---

Last edit 23 July 2022 BJC

```{r setup, include=TRUE, warnings=FALSE, message=FALSE}
knitr::opts_chunk$set(root.dir = "~/OneDrive - James Cook University/Ben PhD/Data & analysis/KimbePreds/", echo = TRUE)
#rm(list=ls())
getwd()
```


# Load relevant packages
```{r packages}
library(readxl)
library(grid)
library(gridExtra)
library(magrittr)
library(ggpubr)
library(sjPlot)
library(cowplot)
library(patchwork)
library(tidyverse)
```



### GRAPHICS FOR SI ###

Need the abun, biom and diversity differences across individual sites/reefs (in addition to reef types)


# ABUNDANCE #

## Load abundance data
```{r abun-load, include=FALSE}
source("../PredAbundance/pred_abun_diw.R", local = knitr::knit_global())
```

## Calculate summary stats for sites
```{r abun-sum-stats}
abunreefstats <- 
  predsum %>% 
  group_by(Reeftype, SiteCode, Site) %>% 
  summarise_each(funs(mean,sd,se=sd(.)/sqrt(n())), No_Preds)

```

## Plot
```{r}
bradabunbar <- 
abunreefstats %>% 
  filter(SiteCode == 'BRAD') %>% 
ggplot(aes(x=SiteCode, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9)) +
  geom_point(aes(shape = Reeftype, fill= Reeftype), colour="black", size = 4) +
  scale_shape_manual(values=c(24, 21, 22)) +                                                       # Kimbe pch scheme
  scale_fill_manual(values=c("#35978F","#436EEE","#DFC27D")) +                        
  scale_y_continuous(limits = c(0,280), "Mean predator count (±SE)") +                             # Need to specify y axis limits as it'll exclude 0 if you don't
  scale_x_discrete("") +                                                                           # Don't need axis names until the bottom panel
  theme_classic() +
  theme(legend.position = "none") +
  #theme(axis.text.x = element_blank()) +
  theme(axis.title.y = element_text(family="Arial", size=10, colour = 'black', vjust=0)) +
  theme(axis.text.y = element_text(family="Arial", size= 10, colour = 'black')) +
  theme(axis.line = element_blank(), panel.border = element_rect(size = 1, fill = "transparent"))

reefabunbar <- 
abunreefstats %>% 
  filter(SiteCode != 'BRAD') %>% 
ggplot(aes(x=SiteCode, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9)) +
  geom_point(aes(shape = Reeftype, fill= Reeftype), colour="black", size = 4) +
  scale_shape_manual(values=c(24, 21, 22)) +                                                       # Kimbe pch scheme
  scale_fill_manual(values=c("#35978F","#436EEE","#DFC27D")) +                        
  scale_y_continuous(limits = c(0,50), name = NULL) +
  scale_x_discrete("") +                                                                           # Don't need axis names until the bottom panel
  theme_classic() +
  theme(legend.position = c(0.8, 0.65)) +
  theme(legend.title=element_blank()) +
  #theme(axis.text.x = element_blank()) +
  theme(axis.title.y = element_text(family="Arial", size=10, colour = 'black', vjust=0)) +
  theme(axis.text.y = element_text(family="Arial", size= 10, colour = 'black')) +
  theme(axis.line = element_blank(), panel.border = element_rect(size = 1, fill = "transparent"))


abunbar <- ggpubr::ggarrange(bradabunbar, reefabunbar, ncol = 2, widths = c(0.2, 1.1) )
abunbar
```



# BIOMASS

```{r biom-load}
source("../PredBiomass/pred_biom_diw.R", local = knitr::knit_global())
```


## Calculate summary stats for sites
```{r abun-sum-stats}
biomreefstats <- 
  merpredsum %>% 
  group_by(Reeftype, SiteCode, Site) %>% 
  summarise_each(funs(mean,sd,se=sd(.)/sqrt(n())), Transect_pred_biomass_kg)
```

## Plot
```{r}
bradbiombar <- 
biomreefstats %>% 
  filter(SiteCode == 'BRAD') %>% 
ggplot(aes(x=SiteCode, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9)) +
  geom_point(aes(shape = Reeftype, fill= Reeftype), colour="black", size = 4) +
  scale_shape_manual(values=c(24, 21, 22)) +                                                       # Kimbe pch scheme
  scale_fill_manual(values=c("#35978F","#436EEE","#DFC27D")) +                        
  scale_y_continuous(limits = c(0,750), "Mean predator biomass (kg ±SE)") +
  scale_x_discrete("") +
  theme_classic() +
  theme(legend.position = "none") + # Don't need axis names until the bottom panel
  #theme(axis.text.x = element_blank()) +
  theme(axis.title.y = element_text(family="Arial", size=10, colour = 'black', vjust=0)) +
  theme(axis.text.y = element_text(family="Arial", size= 10, colour = 'black')) +
  theme(axis.line = element_blank(), panel.border = element_rect(size = 1, fill = "transparent"))


reefbiombar <- 
biomreefstats %>% 
    filter(SiteCode != 'BRAD') %>% 
ggplot(aes(x=SiteCode, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9)) +
  geom_point(aes(shape = Reeftype, fill= Reeftype), colour="black", size = 4) +
  scale_shape_manual(values=c(24, 21, 22)) +                                                       # Kimbe pch scheme
  scale_fill_manual(values=c("#35978F","#436EEE","#DFC27D")) +                        
  scale_y_continuous(limits = c(0,60), NULL) +
  scale_x_discrete("") +
  theme_classic() +
  theme(legend.position = "none") +
  #theme(axis.text.x = element_blank()) +
  theme(axis.title.y = element_text(family="Arial", size=10, colour = 'black', vjust=0)) +
  theme(axis.text.y = element_text(family="Arial", size= 10, colour = 'black')) +
  theme(axis.line = element_blank(), panel.border = element_rect(size = 1, fill = "transparent"))

biombar <- ggpubr::ggarrange(bradbiombar, reefbiombar, ncol = 2, widths = c(0.2, 1.1) )
biombar
```


# DIVERSITY

```{r div-load, warning=FALSE, include=FALSE, echo=TRUE}
source("../PredDiversity/pred_div_diw.R", local = knitr::knit_global())
```


## Calculate summary stats for sites
```{r abun-sum-stats}
divreefstats <- 
  preddiv %>% 
  group_by(Reeftype, SiteCode, Site) %>% 
  summarise_each(funs(mean,sd,se=sd(.)/sqrt(n())), Shannon)
```

```{r}
preddiv %$% 
  summary(SiteCode)
```


## Plot (don't need to have BRAD on separate scale)
```{r}
reefdivbar <- 
divreefstats %>% 
   # filter(SiteCode != 'BRAD') %>% 
ggplot(aes(x=SiteCode, y=mean)) +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.2, position=position_dodge(.9)) +
  geom_point(aes(shape = Reeftype, fill= Reeftype), colour="black", size = 4) +
  scale_shape_manual(values=c(24, 21, 22)) +                                                       # Kimbe pch scheme
  scale_fill_manual(values=c("#35978F","#436EEE","#DFC27D")) +                        
  scale_y_continuous("Mean diversity (H ±SE)") +
  scale_x_discrete("Reef") +
  theme_classic() +
  theme(legend.position = "none") +
  #theme(axis.text.x = element_blank()) +
  theme(axis.title.y = element_text(family="Arial", size=10, colour = 'black', vjust=0)) +
  theme(axis.text.y = element_text(family="Arial", size= 10, colour = 'black')) +
  theme(axis.line = element_blank(), panel.border = element_rect(size = 1, fill = "transparent"))
reefdivbar
```





## And save
```{r all-plot, fig.height=8.5,  fig.width=8.5}
reefbars <- ggpubr::ggarrange(abunbar, biombar, reefdivbar, ncol = 1)
reefbars
ggsave(reefbars, filename= '../../output/rfigures/SI_reefbars.tiff',  width = 174, height = 174, units = "mm", dpi = 600)
ggsave(reefbars, filename= '../../output/rfigures/SI_reefbars.png',  width = 174, height = 174, units = "mm", dpi = 600)
```





